### 2012-06-23 trunk/ijcis-pos-model/sources/temporalModel.tex
more modifications

In this section we will formalize the model for valid-time relational databases. The first subsection is devoted to the formalization of the model. Then, the data manipulation language is defined.

\subsection{\label{subsec:temporal-model}The generalized temporal model}
The model is based on GEFRED~\cite{Medina1994} (Generalized Model of Fuzzy Relational DB) model. This model is extended by adding valid-time support which will be illustrated with examples. The information in the system is defined by the following elements:

@@

\begin{definition}
\label{def:typeof-domain}
\textbf{Typeof(A).}
Consider $D_G$ to be a generalized fuzzy domain and let $A \in D_G$ be a value for the domain. 
The function typeof$\left(A \right)$ returns the datatype associated with the domain $D_G$. For a given value $A$, the function returns the associated number as shown in Table \ref{tbl:gefred-data-types}.
\end{definition}

\vglue13pt
%\begin{table}[htbp]
\tcap{\label{tbl:gefred-data-types}Data types}
%\centerline{\small TITLE}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c p{4cm} }\\
No. & Datatype \\ \hline
1 & A single scalar. \\
2 & A single number. \\
3 & A set of mutually exclusive possible scalar assignations. \\
4 & A set of mutually exclusive possible numeric assignations. \\
5 & A possibility distribution in a scalar domain. \\
6 & A possibility distribution in a numeric domain. \\
7 & A real number in $\left[0, 1 \right]$ referring to degree of matching. \\
8 & An \emph{UNKNOWN} value. \\
9 &  An \emph{UNDEFINED} value. \\
10 & A \emph{NULL} value. \\
\hline\\
\end{tabular}
}

@@

\begin{definition}
\emph{Generalized fuzzy temporal relation.}
A generalized fuzzy temporal relation $R_{FTG}$ is given by~\cite{Medina1994}:
\label{def:generalized-fuzzy-temporal-relation}
\begin{equation}
\label{eq:generalized-fuzzy-temporal-relation}
R_{FTG} = \left(\Head, \Body \right)
\end{equation}
Where $\Head$ is the Head of the relation and consist on a fixed set of triplets attribute- domain - compatibility with an optional the valid-time attribute:

@@

\begin{example}
\label{ex:sample-ft-database}
 Consider a hospital database as explained in example \ref{ex:library-database}. But now, the time is stored with imprecision, because the date of arrival and leave are not precisely known. For simplicity, the ill-known time points are represented as triangular fuzzy numbers. An ill-known time point is given by $[dd/mm/yyyy,a,b]$. The values for $a$ and $b$ are integers. Thus, the date given by $[15/3/2012,5,2]$ is a triangular fuzzy number with the left bound the 10/3/2012, the core on the 15/3/2012 and the right bound on 17/3/2012.
Table \ref{tbl:fuzzy-library-sample} shows the elements Head, $\Head$ and Body, $\Body$. 
\begin{align}
 \label{eq:head-ex}
\Head = & \lbrace \left(\text{ID}:D_{\text{ID}}\right),\\
\nonumber
&  \left(\text{Name}:D_{\text{Name}}\right),\\
% \nonumber
% & \left(\text{Loan}:D_{\text{Loan}}\right) \\
\nonumber
& \left(\text{PVP}:D_{\text{PVP}}\right) \rbrace
\end{align}

@@

\vglue13pt
%\begin{table}[htbp]
\tcap{\label{tbl:fuzzy-library-sample}Sample hospital database}
%\centerline{\small DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c c c |p{0.35\columnwidth}}\\
\hline
& & & \centerline{PVP} \\
$\Head$ &  \textbf{ID}  & Name  &  Arrival , Leave\\
\hline
\multirow{3}{*}{$\Body$ } &  3 & John & [15/3/2012, 5, 2] , [30/3/2012, 1, 1] \\
 &4 & Antoon & [25/3/2012, 3, 2] , [4/4/2012, 1, 7] \\
&5 & Daan & [18/3/2012, 4, 1] , [2/4/2012, 2, 2]\\
% \ 3 & Dracula & [4/4/2012, 3, 3] & UC \\
\hline\\
\end{tabular}
}

@@

For example, in the case of the record with ID = 3:
\begin{align}
%\label{eq:value-component-s}
\nonumber
\Head^{v} &=  \lbrace \left(\text{ID}:D_{\text{ID}}\right), \left(\text{Title}:D_{\text{title}}\right),\\
\nonumber
& \left(\text{PVP}:D_{\text{PVP}}\right) \rbrace \\
\nonumber
\Body^ {v} &= \left \lbrace 3:, \text{``John``}: ,  \left[15/3/2012, 30/3/2012\right] \right \rbrace 
\end{align}

@@

For example, in the case of the record with ID = 3:
\begin{align}
%\label{eq:sample-compatibility-component}
\nonumber
\Head^{c} = & \lbrace \left(C_\text{ID}\right), \left(C_{\text{title}}\right), \left(C_{\text{Loan}}\right), \left(C_{\text{PVP}}\right) \rbrace \\
% \nonumber
% &  \\
% \nonumber
% & \left(C_{\text{Loan}}\right) \\
% \nonumber
% & \left(C_{\text{PVP}}\right) \rbrace \\
\nonumber
\Body^ {c} = &\left \lbrace 1, 1, 1, 1 \right \rbrace 
\end{align}

@@

For example, in the case of the record with ID = 3:
\begin{align}
%\label{eq:sample-temporal-component}
\nonumber
\Head^{t} = & \lbrace \left( \text{PVP}, D_{\text{PVP}}\right),\\
\nonumber
\Body^ {t} =& \left \lbrace 4/4/2012 , UC \right \rbrace 
\end{align}

Analogously, it is possible to define the the value component for the temporal part and the compatibility component for the temporal part.

\begin{definition}
\label{def:generalized-primary-key}
\emph{Generalized primary key.}
Consider $D_G$ to be a fuzzy generalized domain, and let $A_{Gs}:D_{Gs}$ be the attributes and the domain of the attribute for each $s in S \subseteq \left(1, \ldots, n \right)$.
A generalized primary key, $K_G$ is a subset of the head:
\begin{align}
\label{eq:generalized-primary-key}
K_G \subseteq \Head, K_G = \left \lbrace  \left(A_{Gs}:D_{Gs} \right) \right \rbrace \\
\nonumber
s\in S \subseteq \left(1, \ldots, n \right) \\
\nonumber
\forall s \in S, \text{Typeof } \left(D_{Gs} \right) \in \left \lbrace 1, 2 \right \rbrace \\
\nonumber
\forall i, i' \in \left \lbrace 1, \ldots, m\right \rbrace , \exists s \in S: \\
\nonumber
\left(A_{Gs}:d_{is} \right) \neq \left(A_{Gs}:d_{i's} \right)
\end{align}
\end{definition}

@@

\begin{align}
\nonumber
K_G \subseteq \Head, K_G = \left \lbrace  \left(\text{ID}:D_{\text{ID}} \right) \right \rbrace
\end{align}
In this case, the function \emph{Typeof(ID) = 2} (see Definition \ref{def:typeof-domain} and Table \ref{tbl:gefred-data-types}). The primary key for the table is the attribute \emph{ID}, a unique number. Two different books have two different values for the \emph{ID} attribute.

In order to add valid-time support, the primary key should be re-defined. E.g., consider the library database. If the primary key is the \emph{ID} attribute, a book could not be loaned more than once. To resolve this problem, we extend the given primary key with a version identifier. 
%Thus, when a book is loaned, a new entry in the table is added. The \emph{ID} attribute identifies the physical book whereas

\begin{definition}
\label{def:generalized-fuzzy-temporal-key}
\emph{Generalized fuzzy temporal key.}
Consider $D_G$ to be a fuzzy generalized domain, and let $A_{Gs}:D_{Gs}$ be the attributes and the domain of the attribute for each $s in S \subseteq \left(1, \ldots, n \right)$.
A generalized fuzzy temporal key, $K_{GT}$ is a subset of the head. An attribute, namely \emph{version} $V$ is added when dealing with valid-time.
\begin{align}
\label{eq:generalized-fuzzy-temporal-key}
K_{GT} \subseteq \Head, K_{GT} = \left \lbrace  \left(A_{Gs}:D_{Gs} \right) \right.  \\
\nonumber
 \left. \cup  \left(V_{ID}:D_{ID} \right)	\right \rbrace \\
\nonumber
\text{Typeof }\left(D_{ID} \right) = \\
\nonumber
s\in S \subseteq \left(1, \ldots, n \right) \\
\nonumber
\forall s \in S, \text{Typeof } \left(D_{Gs} \right) \in \left \lbrace 1, 2 \right \rbrace \\
\nonumber
\forall i, i' \in \left \lbrace 1, \ldots, m\right \rbrace , \exists s \in S: \\
\nonumber
\left(A_{Gs}:d_{is} \right) \neq \left(A_{Gs}:d_{i's} \right)
\end{align}
\end{definition}

@@

\begin{align}
 \nonumber
K_{GT} \subseteq \Head, K_{GT} = \left \lbrace  \left(\text{ID}:D_{\text{ID}} \right) \cup  \left(V_{\text{ID}}:D_{\text{ID}} \right) \right \rbrace	
%\nonumber
 %\left. \cup  \left(V_{ID}:D_{ID} \right)	\right \rbrace \\
\end{align}

\vglue13pt
%\begin{table}[htbp]
\tcap{\label{tbl:fuzzy-library-sample-valid-time}Sample library database}
%\centerline{\small DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c c c c | p{0.35\columnwidth} }\\
\hline
% & & & \multicolumn{2}{c}{PVP} \\
%& & & & \centerline{PVP}\\
{$\Head$}  &  &  & &(Arrival\\
& \textbf{ID}  & \textbf{V} & Name& ,Leave )	\\
 \hline
\multirow{4}{*}{$\Body$}  & 3 & 001 & John & [15/3/2012, 5, 2], [30/3/2012, 1, 1] \\
& 4 & 001 & Antoon & [25/3/2012, 3, 2] , [4/4/2012, 1, 7]\\
& 5 & 001 & Daan & [18/3/2012, 4, 1] , [2/4/2012, 2, 2]\\
 & 3 & 002 & John & [4/4/2012, 3, 3] , UC \\
\hline\\
\end{tabular}
}

@@

The Generalized Fuzzy Relational Algebra~\cite{Medina1994} manipulates relations like $R_{FG}$. The operations defined are: \emph{Union, Intersection, Difference, Cartesian Product, Projection, Join} and \emph{Selection}. Thus, in this section we will describe and implement the following operations for temporal databases, as described in \ref{subsubsec:Understanding-valid-time-databases}. The operations implemented are: \emph{Insert, Modify, Delete, }. The semantics of the operations will be the same that those defined for a crisp temporal database, whereas the temporal representation is made by the possibilistic valid-time period and the ill-known constraints (see sections \ref{sec:prelim} and \ref{sec:time-rep}.
%The algebra for the data manipulation is defined in .

It is important to notice that while the result of the evaluation of any comparison between crisp time intervals is boolean, the evaluation of any comparison between \emph{PVPs} is in the unit interval.  Therefore, for a valid-time object, say $A \in R$, in a crisp temporal database it is not possible an overlapping among any valid-time interval ($\forall I, \not \exists J: \left(A, I \right) \in R \wedge \left(A, J \right) \in R \wedge \lnot \Overlap\left(I, J \right)$). Thus, from the point of view of a fuzzy temporal database, there are the following options (see Figure \ref{fig:tdb-class}):

% \begin{itemize}
% \item 
\begin{definition}
\textbf{Strictly consistent} It is not possible for two PVPs of the same object to be valid at the same time. In other words, if $A$ is an object, $R$ is a relation and $\pi_{I}\left(t \right)$ and $\pi_{J}\left(t \right)$ are the possibilities for a time point $t \in \T$ to be in the PVPs $I$ and $J$ respectively,
\begin{align}
\label{eq:stricly-consistent}
\forall I,J, \forall t \in \T :\\
%\nonumber
%\left(A, I \right) \in R \wedge \\
%\nonumber
%\left(A, J \right) \in R : \\
\nonumber
\pi_{I}(t) + \pi_{J}(t) \leq 1
\end{align}
\end{definition}

%\item 
\begin{definition}
 \textbf{Ill-consistent} The overlapping of two PVPs of the same object do overlap. It can be distinguished among the following sub-types:
\end{definition}
% 	\begin{itemize}
% 	\item 
	      \begin{definition}
	      \textbf{Co-existence} Two versions of the same object $A$ may exist at the same time.
		      \begin{align}
		      \label{eq:co-existence}
		      \forall I, \exists ! J, \exists t \in \T : \\
		      \nonumber
		      \pi_{I}(t) + \pi_{J}(t) \geq 1
		      \end{align}
	      \end{definition}

	%\item 
	      \begin{definition}
	      \textbf{Weak-consistence} Several versions of the same object $A$ may exist at the same time.
		      \begin{align}
		      \label{eq:weak-consistence}
		      \forall I, \exists J, \exists t \in \T : \\
		      \nonumber
		      \pi_{I}(t) + \pi_{J}(t) \geq 1
		      \end{align}
	      \end{definition}

@@

\vspace*{13pt}
\begin{center}
{
\includegraphics[scale=1.25]{./graphs/tdb-clas.pdf}

}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\fcaption{\label{fig:tdb-class}Classification for a fuzzy temporal database. The validity for an entity is represented by a trapezoid. Graphic (a) shows an example of strictly consistent database. Between two versions of the entity it is possible some amount of overlapping, but always holds equation \eqref{eq:stricly-consistent}. Graphic (b) shows an example of a co-existence. Two versions of the same entity may overlap in any degree, but not more than two versions. Graphic (c) shows an example of weak-consistence. The different versions of the same entity are shown in an unfolded way, to clarify the multiple overlapping.   }
\vspace*{13pt}

@@

Since the time intervals are now possibilistic valid-time periods, PVPs, the auxiliary functions defined in equations \eqref{eq:close-a-crisp-interval} to \eqref{eq:close-current} should be re-defined.

>>>>

In this section we will formalize the model for possibilistic valid-time relational databases. The first subsection is devoted to the formalization of the model. Then, the data manipulation language is defined.

\subsection{\label{subsec:temporal-model}The generalized temporal model}
The model is based on GEFRED~\cite{Medina1994} (Generalized Model of Fuzzy Relational DB) model. This model is extended by adding valid-time support which will be illustrated through definitions and examples. The information in the system is defined by the following elements:

@@

\begin{definition}
\label{def:typeof-domain}
\textbf{Typeof(a).}
Consider $D_G$ to be a generalized fuzzy domain and the elements in definition \ref{def:valid-time-relation}. Let $a$ be the value for the attribute $A$. The function typeof$\left(a \right)$ returns the datatype associated with the value $a$ and returns a number in $\left[1, 10 \right]$ as shown in Table \ref{tbl:gefred-data-types}.
\begin{equation}
 \label{eq:typeof}
\mbox{Typeof} \left(a \right) \mapsto \left[1, 10 \right]
\end{equation}

\begin{samepage}
\vglue13pt
%\begin{table}[htbp]
\tcap{\label{tbl:gefred-data-types}Data types}
%\centerline{\small TITLE}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c p{4cm} }\\
No. & Datatype \\ \hline
1 & A single scalar. \\
2 & A single number. \\
3 & A set of mutually exclusive possible scalar assignations. \\
4 & A set of mutually exclusive possible numeric assignations. \\
5 & A possibility distribution in a scalar domain. \\
6 & A possibility distribution in a numeric domain. \\
7 & A real number in $\left[0, 1 \right]$ referring to degree of matching. \\
8 & An \emph{UNKNOWN} value. \\
9 &  An \emph{UNDEFINED} value. \\
10 & A \emph{NULL} value. \\
\hline\\
\end{tabular}
} 
\end{samepage}

@@

A generalized fuzzy relation is defined in \cite{Medina1994}. Here, we will extend the definition to a generalized fuzzy temporal relation.

\begin{definition}
\emph{Generalized fuzzy temporal relation.}
\label{def:fuzzy-temporal-relation}
Consider the elements in definition \ref{def:valid-time-relation}. Some of the elements will be extended for the fuzzy case.

\begin{itemize}
 \item A set of non-temporal fuzzy  or crisp attributes.
	\begin{equation}
	\label{eq:fuzzy-attribute-set}
	A = \left \lbrace A_1, A_2, \ldots, A_n \right \rbrace
	\end{equation}
      The domain for each attribute $A_1, \ldots, A_n$ is $D_1, \ldots, D_n$ respectively. 
\item The primary key $A_K$ is a subset of $A$.
      \begin{equation}
       \label{eq:fuzzy-primary-key-a}
      A_K \subset A
      \end{equation}
A formal definition of primary key for fuzzy relational databases will be given later in Definition \ref{def:generalized-primary-key}.
\item A set of two attributes; $S$  and $E$ the attributes for the starting and ending ill-known points respectively. $I$ is a possibilistic validity period \emph{PVP} as explained in Section \ref{subsec:ill-known-interval}.
\begin{equation}
 \label{eq:fuzzy-attribute-time-interval}
I = \left \lbrace S, E \right \rbrace
\end{equation}
\item An attribute called version identifier, $V_{ID}$. This attribute is a counter for each different version of the entities. 
% \begin{equation}
%  \label{eq:fuzzy-version-identifier}
% V_{ID} \subset \N
% \end{equation}

\item Then $R_{FTG}$, the schema for the fuzzy valid-time relation is:
\begin{equation}
 \label{eq:fuzzy-valid-time-relation}
R_{FTG} = A \cup V_{ID} \cup  I
\end{equation}
\item The primary key for the fuzzy valid-time relation $R_{FTG}$ is:
\begin{equation}
 \label{eq:fuzzy-valid-time-temporal-pk}
K_{GT} = A_K \cup V_{ID}
\end{equation}
A formal definition of the primary key for fuzzy valid-time relations will be given later in Definition \ref{def:generalized-fuzzy-temporal-key}.

\item We will note by $r$ any valid instance of $R_{FTG}$. 
      \begin{equation}
       \label{eq:fuzzy-valid-time-instance}
      r \subseteq D_1\ x\ \ldots\ x\ D_n
      \end{equation}

\item Let $t$ be a tuple in the instance $r$; $t \in r$:
      \begin{itemize}
      \item We will note the values for the starting and the ending  points,$s$ and $e$  respectively. The value for the time interval is given by $i$.
      \begin{align}
       \label{eq:fuzzy-starting-point}
      s = t\left[S \right]\\
      e = t\left[E \right]\\
      i = \left(s, e\right)
      \end{align}

      \item Let $A_k \subset A$ be the set of the non-temporal attributes that are part of the primary key (equation \eqref{eq:fuzzy-primary-key-a}). Then, $a_k$ denotes the values for the non-temporal attributes of the primary key.
	    \begin{equation}
	     \label{eq:fuzzy-pk-attribute} 
	      a_k = t\left[A_K \right]
	    \end{equation}

    \item Let $K_{GT}$ be the primary key for the valid-time relation as given in equation \eqref{eq:fuzzy-valid-time-temporal-pk}. Then, $k$ denotes the values for the attributes in the primary key.
	  \begin{equation}
	   \label{eq:fuzzy-value-pk}
	  k = t\left[K_{GT} \right]
	  \end{equation}

      \end{itemize}
\item $Vt$ is the set with all the versions for a given tuple $t$.

\begin{equation}
 \label{eq:fuzzy-all-the-versions}
Vt = r\left(t\left[A_K\right] \right)
\end{equation}
If a tuple $t$ contains more than one version, the result of $Vt$ is a set. We will use the index $k$ to address the elements of the set. E.g. $Vt = \left \lbrace t_1, t_2, \ldots, t_n \right \rbrace$. Then $t_k , k \in \left \lbrace 1, \ldots, n \right \rbrace$ is an element of $Vt$.
\end{itemize}
%\end{definition}

We will illustrate the definitions with an example.

\begin{example}
Consider the set of attributes $A = \left \lbrace A_1, A_2, A_3 \right \rbrace$. The primary key for these attributes is given by $A_K = \left \lbrace A_1, A_2 \right \rbrace$. Let $I = \left \lbrace S, E \right \rbrace$ be the set of temporal attributes that define the possibilistic validity period of the data. $R_{FTG} = A \cup V_{ID} \cup I$ is the valid-time relation and $r$ is an instance of the relation. The instance $r$ is given by the following elements. $r = \left \lbrace \left(a_{11}, a_{12}, a_{13}, 001, s_1 ,e_1 \right) \right.$,  $\left(a_{21}, a_{22}, a_{23}, 001, s_2, e_2 \right)$ , $\left. \left(a_{11}, a_{12}, a_{31}, 002, s_3, e_3 \right) \right \rbrace$. The instance $r$ is illustrated in Table \ref{tbl:fuzzy-sample-definitions}. 
Consider the tuple $t = \left(a_{11}, a_{12}, a_{13}, s_1 ,e_1 \right)$. Then,
\begin{align}
 \nonumber
s_1 &= t\left[S \right]\\
 \nonumber
e_1 &= t[E]\\
 \nonumber
i\ \ &= \left(s_1, e_1\right)\\
 \nonumber
a_k &= t\left[A_K\right] = \left(a_{11}, a_{12} \right) \\
 \nonumber
k &= t\left[PK\right] =\left(a_{11}, a_{12}, s_1, e_1\right)\\
 \nonumber
Vt &= r\left(t\left[A_K\right] \right) = \left \lbrace t_1, t_3 \right \rbrace
\end{align}

\begin{samepage}
\vglue13pt
%\begin{table}[htbp]
\tcap{\label{tbl:fuzzy-sample-definitions}Sample database containing the instance $r$ of the fuzzy valid-time relation $R_{FTG}$.}
%\centerline{\small DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c c c c c c c}\\
\hline
& \textbf{A$_1$}  & \textbf{A$_2$}  & $A_3$ & $V_{ID}$ & $S$ & $E$ \\
\hline
$t_1$&$a_{11}$ & $a_{12}$ & $a_{13}$ & $001$ & $s_1$ & $e_1$ \\
$t_2$ & $a_{21}$ & $a_{22}$ & $a_{23}$& $001$ &  $s_2$ & $e_2$ \\
$t_3$ & $a_{11}$ & $a_{12}$ & $a_{31}$& $002$ & $s_3$ & $e_3$\\
\hline\\
\end{tabular}
}
\end{samepage}

% 
% 
% Consider $\A$ to be the set of all the entities, and let $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of (fuzzy) attributes that define an entity. Let $\I_{PVP}$ be the set of all the ill-known time intervals and let $I = \left(X, Y \right)$ be a PVP, $I \in \I_{PVP}$.  The pair $\left(A, I\right)$ expresses that the data regarding the entity A are valid during the ill-known time interval I. Let $R_{FTG} \subseteq \A \  x\  \I_{PVP}$ be a valid-time relation. Then, the following equation indicates that the pair $\left(A, I\right)$ is in the relation R:
% 
% \begin{equation}
% \label{eq:fuzzy-rel-def}
% \left( R_{FTG}, \left(A , I\right) \right)
% \end{equation}
% 
 A generalized fuzzy temporal relation $R_{FTG}$ can be noted also by:
\label{def:generalized-fuzzy-temporal-relation}
\begin{equation}
\label{eq:generalized-fuzzy-temporal-relation}
R_{FTG} = \left(\Head, \Body \right)
\end{equation}
Where $\Head$ is the Head of the relation and consist on a fixed set of triplets attribute- domain - compatibility with an optional the valid-time attribute:

@@

\begin{example}
\label{ex:sample-ft-database}
 Consider a historical database containing diplomatic documents as explained in example \ref{ex:library-database}. But now, the documents are from the Medieval Ages. Hence, the time is known with imprecision. For simplicity, the ill-known time points are represented as triangular fuzzy numbers. An ill-known time point is given by $[dd/mm/yyyy,a,b]$. The values for $a$ and $b$ are integers. Thus, the date given by $[15/3/1012,5,2]$ is a triangular fuzzy number with the left bound the 10/3/1012, the core on the 15/3/1012 and the right bound on 17/3/1012.
Table \ref{tbl:fuzzy-library-sample} shows the elements Head, $\Head$ and Body, $\Body$. 
\begin{align}
 \label{eq:head-ex}
\Head = & \lbrace \left(\text{ID}:D_{\text{ID}}\right),\\
\nonumber
&  \left(\text{Entity}:D_{\text{Entity}}\right),\\
% \nonumber
% & \left(\text{Loan}:D_{\text{Loan}}\right) \\
\nonumber
& \left(\text{PVP}:D_{\text{PVP}}\right) \rbrace
\end{align}

@@

\begin{samepage}
\vglue13pt
%\begin{table}[htbp]
\tcap{\label{tbl:fuzzy-library-sample}Sample historical database}
%\centerline{\small DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c c c |p{0.35\columnwidth}}\\
\hline
& & & \centerline{PVP} \\
$\Head$ &  \textbf{ID}  & Entity  &  Start , End \\
\hline
\multirow{3}{*}{$\Body$ } &  3 & E.U. & [15/3/1012, 5, 2] , [30/3/1012, 1, 1] \\
 &4 & N.A.T.O. & [25/3/1012, 3, 2] , [4/4/1012, 1, 7] \\
&5 & C.E.I. & [18/3/1012, 4, 1] , [2/4/1012, 2, 2]\\
% \ 3 & Dracula & [4/4/1012, 3, 3] & UC \\
\hline\\
\end{tabular}
}
\end{samepage}

@@

For example, in the case of the document with ID = 3:
\begin{align}
%\label{eq:value-component-s}
\nonumber
\Head^{v} &=  \lbrace \left(\text{ID}:D_{\text{ID}}\right), \left(\text{Entity}:D_{\text{entity}}\right),\\
\nonumber
& \left(\text{PVP}:D_{\text{PVP}}\right) \rbrace \\
\nonumber
\Body^ {v} &= \left \lbrace 3:, \text{``E.U.''}: ,  \left[15/3/1012, 30/3/1012\right] \right \rbrace 
\end{align}

@@

For example, in the case of the document with ID = 3:
\begin{align}
%\label{eq:sample-compatibility-component}
\nonumber
\Head^{c} = & \lbrace \left(C_\text{ID}\right), \left(C_{\text{entity}}\right), \left(C_{\text{PVP}}\right) \rbrace \\
% \nonumber
% &  \\
% \nonumber
% & \left(C_{\text{Loan}}\right) \\
% \nonumber
% & \left(C_{\text{PVP}}\right) \rbrace \\
\nonumber
\Body^ {c} = &\left \lbrace 1, 1, 1, 1 \right \rbrace 
\end{align}

@@

For example, in the case of the document with ID = 3:
\begin{align}
%\label{eq:sample-temporal-component}
\nonumber
\Head^{t} = & \lbrace \left( \text{PVP}, D_{\text{PVP}}\right),\\
\nonumber
\Body^ {t} =& \left \lbrace 4/4/1012 , UC \right \rbrace 
\end{align}

Analogously, it is possible to define both the name value component and the compatibility component for the temporal part.

%the value component for the temporal part and the compatibility component for the temporal part.

\begin{definition}
\label{def:generalized-primary-key}
\emph{Generalized primary key.}
Consider $D_G$ to be a fuzzy generalized domain, and let $A_{Gs}:D_{Gs}$ be the attributes and the domain of the attribute for each $s \in S \subseteq \left(1, \ldots, n \right)$.
A generalized primary key, $K_G$ is a subset of the head:
\begin{align}
\label{eq:generalized-primary-key}
K_G \subseteq \Head, K_G = \left \lbrace  \left(A_{Gs}:D_{Gs} \right) \right \rbrace \\
\nonumber
s\in S \subseteq \left(1, \ldots, n \right) \\
\end{align}
Subject to the following constraints:
\begin{align}
%\nonumber
\forall s \in S, \text{Typeof } \left(D_{Gs} \right) \in \left \lbrace 1, 2 \right \rbrace \\
\nonumber
\forall i, i' \in \left \lbrace 1, \ldots, m\right \rbrace , \exists s \in S: \\
\nonumber
\left(A_{Gs}:d_{is} \right) \neq \left(A_{Gs}:d_{i's} \right)
\end{align}
\end{definition}

@@

\begin{align}
\nonumber
K_G \subseteq \Head, K_G = \left \lbrace  \left(\text{ID}:D_{\text{ID}} \right) \right \rbrace
\end{align}
In this case, the function \emph{Typeof(ID) = 2} (see Definition \ref{def:typeof-domain} and Table \ref{tbl:gefred-data-types}). The primary key for the table is the attribute \emph{ID}, a unique number. Two different documents have two different values for the \emph{ID} attribute.

In order to add valid-time support, the primary key should be re-defined. E.g., consider the historical database. If the primary key is the \emph{ID} attribute, a document should be valid only during one period of time. To resolve this problem, we extend the given primary key with a version identifier.

%a book could not be loaned more than once. 
%Thus, when a book is loaned, a new entry in the table is added. The \emph{ID} attribute identifies the physical book whereas

\begin{definition}
\label{def:generalized-fuzzy-temporal-key}
\emph{Generalized fuzzy temporal key.}
Consider $D_G$ to be a fuzzy generalized domain, and let $A_{Gs}:D_{Gs}$ be the attributes and the domain of the attribute for each $s \in S \subseteq \left(1, \ldots, n \right)$. Let  $V$ be a new attribute called \emph{version}. 
A generalized fuzzy temporal key, $K_{GT}$ is a subset of the head.
\begin{align}
\label{eq:generalized-fuzzy-temporal-key}
K_{GT} \subseteq \Head, K_{GT} = \left \lbrace  \left(A_{Gs}:D_{Gs} \right) \right.  \\
\nonumber
 \left. \cup  \left(V_{ID}:D_{ID} \right)	\right \rbrace \\
\nonumber
\text{Typeof }\left(D_{ID} \right) = \\
\nonumber
s\in S \subseteq \left(1, \ldots, n \right) \\
\end{align}
Subject to the following constraints:
\begin{align}
\forall s \in S, \text{Typeof } \left(D_{Gs} \right) \in \left \lbrace 1, 2 \right \rbrace \\
\nonumber
\forall i, i' \in \left \lbrace 1, \ldots, m\right \rbrace , \exists s \in S: \\
\nonumber
\left(A_{Gs}:d_{is} \right) \neq \left(A_{Gs}:d_{i's} \right)
\end{align}
\end{definition}

@@

\begin{align}
 \nonumber
K_{GT} \subseteq \Head, K_{GT} = \left \lbrace  \left(\text{ID}:D_{\text{ID}} \right) , \left(V_{\text{ID}}:D_{\text{ID}} \right) \right \rbrace	
%\nonumber
 %\left. \cup  \left(V_{ID}:D_{ID} \right)	\right \rbrace \\
\end{align}

\begin{samepage}
\vglue13pt
%\begin{table}[htbp]
\tcap{\label{tbl:fuzzy-library-sample-valid-time}Sample historical database}
%\centerline{\small DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c c c c | p{0.35\columnwidth} }\\
\hline
% & & & \multicolumn{2}{c}{PVP} \\
%& & & & \centerline{PVP}\\
{$\Head$}  &  &  & &(Start \\
& \textbf{ID}  & \textbf{V} & Entity& ,End )	\\
 \hline
\multirow{4}{*}{$\Body$}  & 3 & 001 & E.U. & [15/3/1012, 5, 2], [30/3/1012, 1, 1] \\
& 4 & 001 & N.A.T.O. & [25/3/1012, 3, 2] , [4/4/1012, 1, 7]\\
& 5 & 001 & C.E.I. & [18/3/1012, 4, 1] , [2/4/1012, 2, 2]\\
 & 3 & 002 & E.U. & [4/4/1012, 3, 3] , UC \\
\hline\\
\end{tabular}
}
\end{samepage}

@@

The Generalized Fuzzy Relational Algebra~\cite{Medina1994} manipulates relations like $R_{FTG}$. The operations defined are: \emph{Union, Intersection, Difference, Cartesian Product, Projection, Join} and \emph{Selection}. Thus, in this section we will describe and implement the following operations for temporal databases, as described in \ref{subsubsec:Understanding-valid-time-databases}. The operations implemented are: \emph{Insert, Modify} and \emph{Delete}. The semantics of the operations will be the same that those defined for a crisp temporal database, whereas the temporal representation is made by the possibilistic valid-time period and the ill-known constraints (see sections \ref{sec:prelim} and \ref{sec:time-rep}.
%The algebra for the data manipulation is defined in .

It is important to notice that while the result of the evaluation of any comparison between crisp time intervals is boolean, the evaluation of any comparison between \emph{PVPs} is a value in the unit interval. Therefore, for a crisp valid-time relation, it is not possible an overlapping among any valid-time interval. In other words, a crisp temporal database is always strictly consistent. Thus, from the point of view of a fuzzy temporal database, there are the following choices (see Figure \ref{fig:tdb-class}):

% \begin{itemize}
% \item 
\begin{definition}
\textbf{Strictly consistent} It is not possible that two versions of an entity to be valid at the same time.
Consider a tuple $t$ in the instance $r$ of the fuzzy valid-time relation $R_{FTG}$ and  $Vt = \left \lbrace t_1, \ldots, t_n \right \rbrace$ as defined in \ref{def:fuzzy-temporal-relation}. Then, a fuzzy temporal database is strictly consistent if: 
\begin{align}
\label{eq:stricly-consistent}
\forall k,j \in \left \lbrace 1, \ldots, n \right \rbrace, k < j\\
\nonumber t_k \mbox{ Overlaps } t_j \leq 1
\end{align}
\end{definition}

%\item 
\begin{definition}
 \textbf{Ill-consistent} The overlapping of two PVPs of the same entity do overlap. It can be distinguished among the following sub-types:
\end{definition}
% 	\begin{itemize}
% 	\item 
	      \begin{definition}
	      \textbf{Co-existence} Two versions of the same entity may exist at the same time.
		      \begin{align}
		      \forall k, \exists !j \in \left \lbrace 1, \ldots, n \right \rbrace, k < j\\
		      \nonumber t_k \mbox{ Overlaps } t_j \geq 1
		      \end{align}
	      \end{definition}

	%\item 
	      \begin{definition}
	      \textbf{Weak-consistence} Several versions of the same entity may exist at the same time.
		      \begin{align}
		      \label{eq:weak-consistence}
		      \forall k, \exists j \in \left \lbrace 1, \ldots, n \right \rbrace, k < j\\
		      \nonumber t_k \mbox{ Overlaps } t_j \geq 1
		      \end{align}
	      \end{definition}

@@

\begin{samepage}
\vspace*{13pt}
\begin{center}
{
\includegraphics[scale=1.25]{./graphs/tdb-clas.pdf}

}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\fcaption{\label{fig:tdb-class}Classification for a fuzzy temporal database. The validity for an entity is represented by a trapezoid. Graphic (a) shows an example of strictly consistent database. Between two versions of the entity it is possible some amount of overlapping, but always holds equation \eqref{eq:stricly-consistent}. Graphic (b) shows an example of a co-existence. Two versions of the same entity may overlap in any degree, but not more than two versions. Graphic (c) shows an example of weak-consistence. The different versions of the same entity are shown in an unfolded way, to clarify the multiple overlapping.   }
\vspace*{13pt}
\end{samepage}

@@

Since the time intervals are now possibilistic valid-time periods, PVPs, the auxiliary functions defined in equations \eqref{eq:close-a-crisp-interval} to \eqref{eq:close-current} are the basis for the following auxiliary functions.
